\subsection{Graphing the Results}
\label{graphing}

In \texttt{cplint} on SWISH you can draw graphs
for visualizing the results either with \href{http://www.c3js.org/}{C3.js} or with \href{https://www.r-project.org/}{R}. Similar predicates are avaiiable for the two methods.
 There are two types
of graphs: those that represent individual probability values with a bar chart and those that
visualize the results of sampling arguments.

\subsubsection{Using C3.js}
You can draw the probability of a query being true and
being false as a bar chart using the predicates
\begin{verbatim}
bar1(+Probability:float,-Chart:dict) is nondet
bar(+Probability:float,-Chart:dict) is nondet
bar(+Successes:int,+Failures:int,-Chart:dict) is nondet
argbar(+Values:list,-Chart:dict) is nondet
\end{verbatim}
They return  a dict for rendering with C3.js as a bar chart:
the first  returns bar chart with
a single bar for the probability,  the second a chart with
bar for the probability and a bar for one minus the probability,
the third a  chart with
a bar for the number of successes and a bar for the number of failures, and
the fourth a  chart with
a bar each value, where \verb|Values| is a list of couples \verb|V-N| where
  \verb|V| is the value and \verb|N| is the number of samples
  returning that value.
 
To render C3.js charts  you have to include
\begin{verbatim}
:- use_rendering(c3).
\end{verbatim}
before \verb|:- pita.| 

You can also use the  \verb|bar(-Chart:dict)| option of many predicates
as in
\begin{verbatim}
?- prob(heads(coin),biased(coin),P,[bar(Chart)]).
\end{verbatim}
\verb|P| will be instantiated with a
 chart with
a bar for the probability of \verb|heads(coin)| true and a bar for the probability of \verb|heads(coin)| false,
given that \verb|biased(coin)| is true.

Another example is
\begin{verbatim}
?- mc_prob(heads(coin),P,[bar(Chart)]).
\end{verbatim}
that returns a chart representation of the probability.
\begin{verbatim}
?- mc_sample(heads(coin),1000,P,[bar(Chart)]).
\end{verbatim}
returns in \verb|Chart| a diagram with one bar for the number of successes and
one bar for the number of failures.

The options of
\verb|mc_sample_arg/5|, \verb|mc_sample_arg_first/5|,   \verb|mc_mh_sample_arg/6|,  \verb|mc_rejection_sample_arg/6|, 
can be used for visualizing the results of sampling arguments.

An example is
\begin{verbatim}
?- mc_sample_arg(reach(s0,0,S),50,S,ValList,[bar(Chart)]).
\end{verbatim}
of \href{http://cplint.eu/example/inference/markov_chain.pl}{\texttt{markov\_chain.pl}}.

The same result can be achieved with
\begin{verbatim}
?- mc_sample_arg(reach(s0,0,S),50,S,ValList),argbar(ValList,Chart)
\end{verbatim}
Drawing a graph is particularly interesting when
sampling values for continuous arguments of goals.
In this case, you can use the samples to draw the
probability density function of the argument.
The predicate
\begin{verbatim}
histogram(+List:list,-Chart:dict,+Options:list)
\end{verbatim}
draws a histogram of the samples in \verb|List| dividing the domain in
 \verb|NBins| bins.
\verb|List| must be a list of couples of the form \verb|[V]-W| or  \verb|V-W|
where \verb|V| is a sampled value and \verb|W| is its weight. This is the format of the list of samples returned by argument sampling predicates.

The predicate
\begin{verbatim}
density(+List:list,-Chart:dict,+Options:list)
\end{verbatim}
draws a line chart of the density the samples in \verb|List| dividing the domain in
 \verb|NBins| bins.
\verb|List| must be as for \verb|histogram/4|.

In  \verb|histogram/4| and  \verb|density/4| \verb|Options| is a list of options, the following are recognised: \begin{itemize}
\item \verb|min(+Min:float)|
the minimum value of domain, default value the minimum in \verb|List|
\item \verb|max(+Max:float)|
the maximum value of domain, default value the maximum in  \verb|List|
\item \verb|nbins(+NBins:int)|
  the number of bins for dividing the domain, default value 40
\end{itemize}
In this way you can specify the limits and the number of intervals of the $X$.


The predicate
\begin{verbatim}
densities(+PriorList:list,+PostList:list,-Chart:dict,+Options:list)
\end{verbatim}
draws a line chart of the density of two sets of samples, usually
 prior and post observations. The samples in \verb|PriorList| and \verb|PostList|
can be either couples \verb|[V]-W| or \verb|V-W| where \verb|V| is a value and \verb|W| its weight.
The same options as for \verb|histogram/4| and  \verb|density/4|  are recognized.

For example
\begin{verbatim}
?-  mc_sample_arg(value(0,X),1000,X,L0,[]),
    histogram(L0,Chart,[]).
\end{verbatim}
from \href{http://cplint.eu/example/inference/gauss_mean_est.pl}{\texttt{gauss\_mean\_est.pl}}
takes 1000 samples of argument \verb|X| of \verb|value(0,X)| and draws the density of the samples using an histogram.

Instead
\begin{verbatim}
?- mc_sample_arg(value(0,Y),1000,Y,L0,[]),
   mc_lw_sample_arg(value(0,X),
    (value(1,9),value(2,8)),1000,X,L),
   densities(L0,L,Chart).
\end{verbatim}
from \href{http://cplint.eu/example/inference/gauss_mean_est.pl}{\texttt{gauss\_mean\_est.pl}}
takes 1000 amples of argument \verb|X| of \verb|value(0,X)| before and after observing
\verb|(value(1,9),value(2,8)| and draws the prior and posterior densities of the samples using a line chart.

You can obtain the BDD for a query with the predicate
\begin{verbatim}
bdd_dot_string(:Query:atom,-DotString:string,-Var:list)
\end{verbatim}
if you include
\begin{verbatim}
:- use_rendering(graphviz).
\end{verbatim}
before \verb|:- pita.|
Solid edges indicate 1-children, dashed edges indicate 0-children and dotted
edges indicate 0-children with negation applied to the sub BDD.
Each level of the BDD is associated to a variable of the form XI\_J indicated on the left:
I indicates the  multivalued variable index and J the index of the Boolean variable of rule I.
The hexadecimal number in each node is part of its address in memory and is not significant.
The table \verb|Var| contains the associations between the rule groundings and the
multivalued variables.

For example (\href{http://cplint.eu/example/inference/coin.pl}{\texttt{coin.pl}})
\begin{verbatim}
?- bdd_dot_string(heads(coin),BDD,Var).
\end{verbatim}
returns the BDD for the query \verb|heads(coin)| and the list of associations between rule groundings and
multivalued variables.

\subsubsection{Using R}
You have to load library \texttt{cplint\_r}  (a SWI-Prolog pack) with
\begin{verbatim}
:- use_module(library(cplint_r)).
\end{verbatim}
Then you can use predicates
\begin{verbatim}
prob_bar_r/1, prob_bar_r/2
mc_prob_bar_r/3
mc_sample_bar_r/2
mc_sample_arg_bar_r/3
mc_sample_arg_first_bar_r/3
mc_rejection_sample_arg_bar_r/4
mc_mh_sample_arg_bar_r/5
mc_mh_sample_arg_bar_r/6
\end{verbatim}

You also have
\begin{verbatim}
histogram_r(+List:list,+NBins:int) is det
\end{verbatim}
that works as \texttt{histogram/3} but does not return the graph as an argument as the graph is
printed with a different mechanism.
\begin{verbatim}
density_r(+List:list) is det
\end{verbatim}
is like \texttt{density/3} but does not require the number of bins in input, they are determined
by R.
\begin{verbatim}
densities_r(+PriorList:list,+PostList:list) is det
\end{verbatim}
is like \texttt{densities/4} but does not require the number of bins in input.

See \url{https://frnmst.github.io/cplint_r/cplint_r.html} for a more detailed manual.

See \href{http://cplint.eu/example/inference/gauss_mean_est_R.pl}{\texttt{gauss\_mean\_est\_R.pl}} for an example of use of these predicates.


